---
title: "Rural Credit and Deforestation"
author: "George Porto Ferreira"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r Setup, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!"tidyverse" %in% installed.packages()) {
  install.packages("tidyverse")
}
if (!"fs" %in% installed.packages()) {
  install.packages("fs")
}
if (!"here" %in% installed.packages()) {
  install.packages("here")
}
if (!"tsibble" %in% installed.packages()) {
  install.packages("tsibble")
}
if (!"terra" %in% installed.packages()) {
  install.packages("terra")
}

library(tidyverse)
library(fs)
library(here)
library(tsibble)
library(terra)

options(timeout = 5500) # (in seconds) if your are in a limited connection increase timeout
states <- c("RO", "PA", "MT")
```

# SICOR

Rural Credit data are made available by the Central Bank of Brazil from
SICOR Database. More details about the types of credit and other
information can be accessed through the link
<https://www.bcb.gov.br/estabilidadefinanceira/creditorural>

In this work we accessed data from three tables with data on rural
properties and loan amounts received (SICOR_PROPRIEDADES, SICOR_OPERACAO_BASICA and
SICOR_LIBERACAO_RECURSOS)

Model and data dictionary are available at the link below:
<https://www.bcb.gov.br/estabilidadefinanceira/creditorural?modalAberto=tabelas_sicor>

The code below loads the data from these three tables into memory.
If the tables do not exist in the working directory, they will be downloaded.

The code below loads the data from these three tables into a Spatialite Database named rural_credit for visualization using QGIS GIS Desktop. 


```{r SICOR Data, eval=FALSE}
# SICOR_PROPRIEDADES --------------------
# https://www.bcb.gov.br/htms/sicor/DadosBrutos/SICOR_PROPRIEDADES.gz
url_sicor <- "https://www.bcb.gov.br/htms/sicor/DadosBrutos/"

if (!file.exists(here::here("SICOR_PROPRIEDADES.gz"))) {
  download.file(paste0(url_sicor, "SICOR_PROPRIEDADES.gz"), here::here("SICOR_PROPRIEDADES.gz"))
}

sicor_propriedades <- gzfile(here::here("SICOR_PROPRIEDADES.gz")) |>
  read.csv(sep = ";") |>
  mutate(
    REF_BACEN = X.REF_BACEN,
    cod_imovel = paste0(substr(CD_CAR, 1, 2), "-", substr(CD_CAR, 3, 9), "-", substr(CD_CAR, 10, 41)),
    uf = paste0(substr(CD_CAR, 1, 2))
  ) |>
  dplyr::select(-X.REF_BACEN, -CD_CAR, -CD_CNPJ_CPF, -CD_SNCR, -CD_NIRF) |>
  dplyr::filter(uf %in% states)

# SICOR LIBERACAO RECURSOS --------------------------
if (!file.exists(here::here("SICOR_LIBERACAO_RECURSOS.gz"))) {
  download.file(paste0(url_sicor, "SICOR_LIBERACAO_RECURSOS.gz"), here::here("SICOR_LIBERACAO_RECURSOS.gz"))
}

sicor_liberacao_recursos <- gzfile(here::here("SICOR_LIBERACAO_RECURSOS.gz")) |> # 2min
  read.csv(sep = ";") |>
  mutate(year_treat = as.integer(substr(X.LIR_DT_LIBERACAO, 7, 10)), DT_LIBERACAO = X.LIR_DT_LIBERACAO) |>
  dplyr::select(-X.LIR_DT_LIBERACAO) |>
  dplyr::filter(year_treat > 2012 & year_treat < 2024) |>
  group_by(REF_BACEN, NU_ORDEM, year_treat) |> # There are different values for same REF_BACEN and NU_ORDEM
  summarize(VL_LIBERADO = sum(LIR_VL_LIBERADO), .groups = "drop")

# SICOR_OPERACAO_BASICA --------------------------
# https://www.bcb.gov.br/htms/sicor/DadosBrutos/SICOR_OPERACAO_BASICA *YEAR* .gz

sicor_operacao_basica <- list()
# Function to check if the file exists and read or download-and-read the file
process_year <- function(ano) {
  file_path <- here::here(paste0("SICOR_OPERACAO_BASICA_ESTADO_", ano, ".gz"))
  if (!file.exists(file_path)) {
    download.file(paste0(url_sicor, "SICOR_OPERACAO_BASICA_ESTADO_", ano, ".gz"), file_path)
  }
  data <- read.csv(gzfile(file_path), sep = ";") %>%
    mutate(
      REF_BACEN = X.REF_BACEN, # create ID
      DT_EMISSAO = as.Date(DT_EMISSAO, format = "%m/%d/%Y"),
      DT_VENCIMENTO = as.Date(DT_VENCIMENTO, format = "%m/%d/%Y")
    ) |>
    dplyr::select(-CD_CONTRATO_STN, -X.REF_BACEN) |> # Drop unused columns
    dplyr::filter(CD_ESTADO %in% states)
  return(data)
}

years <- 2013:(year(Sys.Date()) - 1)
sicor_operacao_basica <- map(years, process_year, .progress = TRUE) %>% bind_rows()
```

# CAR

Spatial data with the delimitation of rural properties are made
available by the Brazilian Forestry Service. Public access offers only
the manual download state by state through the link
<https://www.car.gov.br/publico/estados/downloads>

The Legal Amazon comprises 9 Brazilian states. Amazonas, Para, Mato
Grosso, Amapá, Roraima, Rondônia, Acre, Totantins and half of Maranhão,

Before running the code below, the shapefiles for each state must be
downloaded and stored in different folders.

The code below concatenates the shapefiles from the nine states.

```{r Create Folder CAR}
# Define the path to the folder
folder_path <- paste0("C:/Users/", Sys.getenv("USERNAME"), "/Downloads/car")

# Check if the folder already exists
if (!dir.exists(folder_path)) {
  # Create the folder if it does not exist
  dir.create(folder_path)
  cat("Folder 'CAR' created successfully.\n")
} else {
  cat("Folder 'CAR' already exists.\n")
}

ext <- c(".dbf", ".shp", ".shx", ".prj")
```

Once the folder is created, compacted shapefiles can now be manually downloaded from <https://www.car.gov.br/publico/estados/downloads>, click on "Base de Downloads" and download desired states to `r folder_path`

```{r Download CAR manually, eval=FALSE}
# Download manually since there is a human check to avoid bots
```

Run the chunk below to unzip files

```{r Unzip Files, eval=FALSE}
zipfiles <- folder_path %>% dir_ls()

for (i in 1:length(zipfiles)) {
  unzip(zipfiles[i], exdir = substr(zipfiles[i], 1, nchar(zipfiles[i]) - 4))
}

level1_unziped_folders <- substr(zipfiles, 1, nchar(zipfiles) - 4) %>%
  dir_ls() %>%
  str_subset("AREA_IMOVEL_1.shp$")

for (i in 1:length(level1_unziped_folders)) {
  file.rename(
    paste0(substr(zipfiles[i], 1, nchar(zipfiles[i]) - 4), "/AREA_IMOVEL_1", ext),
    paste0(
      substr(zipfiles[i], 1, nchar(zipfiles[i]) - 4),
      substr(zipfiles[i], nchar(zipfiles[i]), nchar(zipfiles[i]) - 4), ext
    )
  )
}
```

Create CAR SaptialVect from three states

```{r Load and Merge CAR Shapefiles, eval=FALSE} 
# Getting all file paths
folder_path <- paste0("C:/Users/", Sys.getenv("USERNAME"), "/Downloads/car")
shapefiles_car <- folder_path |> 
  dir_ls(recurse = TRUE) |> 
  str_subset(".shp$")

car_aml <- shapefiles_car |> 
  map(vect, crs = "epsg:4674", .progress = TRUE) |> 
  reduce(rbind) |> 
  terra::subset(
    des_condic != "Cancelado por duplicidade" & # duplicates
      ind_tipo == "IRU" & # IRU = Rural Properties
      num_area >= 1 &  # keep only properties grater than 1 ha
    ind_status %in% c("AT","PE"),
     c(cod_imovel, num_area, municipio, cod_estado), 
    NSE = TRUE
  )

# Remove remaining duplicated cod_imovel based on rural property size (keep larger properties)
car_aml_df <- as.data.frame(car_aml)
car_aml_df$wkt <- car_aml@ptr$getGeometryWKT()

car_aml_df <- car_aml_df |> dplyr::group_by(cod_imovel) |> 
  dplyr::slice_max(num_area, with_ties = FALSE) |> 
  dplyr::ungroup() |> 
  filter(cod_estado %in% states)

car_aml <- vect(car_aml_df, crs = "epsg:4674", geom="wkt") |>
    makeValid()



```
# INPE

Every year the National Institute for Space Research (INPE) maps out the
increase in deforestation based on satellite images. Information about
the PRODES Project can be accessed through the link
<http://www.obt.inpe.br/OBT/assuntos/programas/amazonia/prodes> The
methodology @shimabukuroMappingMonitoringDeforestation2000
@valerianoMonitoringTropicalForest2004 used since 1988 has been improved
over the years and today all policies to combat deforestation and the
Brazilian NDC use PRODES data as a basis.



The code below loads the data for processing.

```{r Deforestation Data PRODES, eval=FALSE}
# Deforestation ----
inpe_path <- paste0("C:/Users/", Sys.getenv("USERNAME"), "/Downloads/")

prodes_url <- "https://terrabrasilis.dpi.inpe.br/download/dataset/legal-amz-prodes/vector/yearly_deforestation.zip"

# Define paths
prodes_path <- paste0(inpe_path, "yearly_deforestation.shp")
zipfile_path <- paste0(inpe_path, "yearly_deforestation.zip")

# Function to download and unzip the shapefile
download_and_unzip <- function(url, zip_path, exdir) {
  download.file(url, zip_path)
  unzip(zip_path, exdir = exdir)
  file.remove(zip_path)
}

# Check if the shapefile exists
if (!file.exists(prodes_path)) {
  download_and_unzip(prodes_url, zipfile_path, inpe_path)
}

prodes <- vect(prodes_path, crs = "EPSG:4674") %>%
  makeValid() %>%
  subset(state %in% states, NSE = TRUE) # 50sec

```

# Analysis

The analysis starts by identifying which rural properties suffered
deforestation after August 2008. It is important to understand that the
PRODES data always comprise a 12-month period from August of one year to
July of the following year.

Another important concept to be understood is that of the expiration of
the punitive intention. According to law 9605, in general, deforestation
that has been committed 5 years or more ago can no longer receive a
fine. In addition, law 12651 determines that deforestation committed
without authorization after July 22, 2008 is not eligible for
regularization.

For these reasons, only data from PRODES after July 2008 are being used
in this analysis. Furthermore, the SICOR data begin in 2013, which
allows us to verify if, after 2013, any properties that received rural
credit had committed deforestation in the previous 5 years.

```{r Calculations, eval=FALSE}

{deforestation_in_car <- prodes * car_aml #1min
Sys.time()}

deforestation_in_car$area_deforestation <- expanse(deforestation_in_car, unit = "ha")

deforestation_in_car_df <- as.data.frame(deforestation_in_car) %>%
  group_by(cod_imovel, year) %>%
  summarise(
    deforestation_ha = as.numeric(sum(area_deforestation, na.rm = TRUE)),
    num_area = first(num_area),
    municipio = first(municipio),
    cod_estado = first(cod_estado),
    .groups = "drop"
  ) %>%
  mutate(
    percent_deforestation = as.numeric(deforestation_ha / num_area),
    year_prodes = year
  ) %>%
  select(-year)


# WARNING! There will be duplicated values for deforestation area inherited from deforestation_in_car and also duplicated values in credit values because cardinality between sicor_liberacao_recursos and sicor_propriedades is many-to-many. The dataframe below should only be analysed at the rural property level. Deforestation data should not be aggregated because there are overlapping rural properties.
{
deforestation_and_values <- full_join(deforestation_in_car_df, #1min
  sicor_propriedades,
  by = "cod_imovel",
  relationship = "many-to-many"
) %>%
  group_by(cod_imovel) %>%
  summarise(
    deforestation_ha = sum(deforestation_ha),
    municipio = first(municipio),
    cod_estado = first(cod_estado),
    num_area = first(num_area),
    REF_BACEN = first(REF_BACEN),
    NU_ORDEM = first(NU_ORDEM),
    .groups = "drop"
  ) %>%
  full_join(sicor_liberacao_recursos,
    by = c("REF_BACEN", "NU_ORDEM"),
    relationship = "many-to-many"
  ) %>%
  group_by(cod_imovel) %>%
  summarise(
    deforestation_ha = first(deforestation_ha),
    municipio = first(municipio),
    cod_estado = first(cod_estado),
    vl_liberado = sum(VL_LIBERADO),
    .groups = "drop"
  )
Sys.time()
}

deforestation_and_values_filtered <- deforestation_and_values %>%
  dplyr::filter(!is.na(deforestation_ha) &
    (!is.na(vl_liberado)))

#Treat CAR overlapping

car_w_rc_and_def <- filter(deforestation_and_values, vl_liberado > 0 & deforestation_ha > 0)
car_w_rc_and_def <- car_aml[car_aml$cod_imovel %in% car_w_rc_and_def$cod_imovel, ] |> 
  terra::aggregate(by = "cod_estado")

prodes_dissolve <- terra::aggregate(prodes) #15min
deforestation_in_car_w_rc <- prodes_dissolve * car_w_rc_and_def #1min

deforestation_in_car_w_rc$area_deforestation <- expanse(deforestation_in_car_w_rc, unit = "ha")

deforestation_in_car_w_rc_df <- as.data.frame(deforestation_in_car_w_rc)

```
# Statistics

```{r Statistics, eval=FALSE}
#

results <- deforestation_and_values %>%
  dplyr::filter(!is.na(deforestation_ha) &
    (!is.na(vl_liberado))) %>%
  group_by(cod_estado) %>%
  summarise(
    vl_liberado = sum(vl_liberado),
    qtd_property = n(),
    .groups = "drop"
  ) |> left_join(deforestation_in_car_w_rc_df) |> 
  select(-cod_imovel, -municipio, -agg_n)

results
  
```

# Results

Statistics for total amount of Deforestation, Rural Properties and Rural Credit per state came from the following sources:

Deforestation: INPE. (2024). Incremento no desmatamento da Amazônia Legal à partir de 2008 (Geospatial a5220c18-f7fa-4e3e-b39b-feeb3ccc4830; Version 2023) [ESRI-ShapeFile]. Terrabrasilis. https://terrabrasilis.dpi.inpe.br/geonetwork/srv/eng/catalog.search#/metadata/a5220c18-f7fa-4e3e-b39b-feeb3ccc4830

Rural Properties: SFB. (2024). Cadastro Ambiental Rural (Version v3.47.1) [Geospatial; ESRI-ShapeFile]. SICAR. https://consultapublica.car.gov.br/publico/imoveis/index

Rural Credit:BCB. (2024). Tabelas e Microdados do Crédito Rural e do Proagro (Version 2024) [Table; CSV]. SICOR. https://www.bcb.gov.br/estabilidadefinanceira/tabelas-credito-rural-proagro

# Spatialite Database creation---

```{r, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!"RSQLite" %in% installed.packages()) {
  install.packages("RSQLite")
}
if (!"DBI" %in% installed.packages()) {
  install.packages("DBI")
}
if (!"archive" %in% installed.packages()) {
  install.packages("archive")
}

library(RSQLite)
library(DBI)
library(archive)
```

The code below creates a Spatialite Database named rural_credit to explore spatial data using QGIS Desktop. 
```{r Spatialite download, eval=FALSE}
# Download and create Spatialite DataBase ----
url_spatialite <- "http://www.gaia-gis.it/gaia-sins/windows-bin-amd64/mod_spatialite-5.1.0-win-amd64.7z"
if (!file.exists(here::here("mod_spatialite-5.1.0-win-amd64.7z"))) {
  download.file(url_spatialite, here::here("mod_spatialite-5.1.0-win-amd64.7z"))
}

file <- "mod_spatialite-5.1.0-win-amd64.7z"
archive_extract(file)

# Get the current PATH environment variable
current_path <- Sys.getenv("PATH")

spatialite_path <- here::here("mod_spatialite-5.1.0-win-amd64")
# Add the new path to the existing PATH
new_path <- paste(current_path, spatialite_path, sep = ";")

# Set the new PATH environment variable
Sys.setenv(PATH = new_path)
```

# Creat connection to Spatialite DataBase

```{r Connect, eval=FALSE}
# Create a connection to a new SQLite database file where sicor tables will be stored
paste("Database will be created at", getwd())

conn <- dbConnect(RSQLite::SQLite(), "rural_credit.sqlite")

# Enable SpatiaLite extension
dbExecute(conn, "SELECT load_extension('mod_spatialite');")

# Initialize SpatiaLite metadata (it will take several minutes)
dbExecute(conn, "SELECT InitSpatialMetaData();")
```

# Load data to Spatialite database

```{r Creat tables, eval=FALSE}
# Write sicor_operacao_basica data to DB ----
dbExecute(
  conn,
  "CREATE TABLE `sicor_operacao_basica` (
  `REF_BACEN` INTEGER,
  `NU_ORDEM` INTEGER,
  `CNPJ_IF` INTEGER,
  `DT_EMISSAO` REAL,
  `DT_VENCIMENTO` REAL,
  `CD_INST_CREDITO` INTEGER,
  `CD_CATEG_EMITENTE` INTEGER,
  `CD_FONTE_RECURSO` INTEGER,
  `CNPJ_AGENTE_INVEST` INTEGER,
  `CD_ESTADO` TEXT,
  `CD_REF_BACEN_INVESTIMENTO` REAL,
  `CD_TIPO_SEGURO` INTEGER,
  `CD_EMPREENDIMENTO` REAL,
  `CD_PROGRAMA` INTEGER,
  `CD_TIPO_ENCARG_FINANC` INTEGER,
  `CD_TIPO_IRRIGACAO` INTEGER,
  `CD_TIPO_AGRICULTURA` INTEGER,
  `CD_FASE_CICLO_PRODUCAO` INTEGER,
  `CD_TIPO_CULTIVO` INTEGER,
  `CD_TIPO_INTGR_CONSOR` INTEGER,
  `CD_TIPO_GRAO_SEMENTE` INTEGER,
  `VL_ALIQ_PROAGRO` REAL,
  `VL_JUROS` REAL,
  `VL_PRESTACAO_INVESTIMENTO` REAL,
  `VL_PREV_PROD` REAL,
  `VL_QUANTIDADE` REAL,
  `VL_RECEITA_BRUTA_ESPERADA` REAL,
  `VL_PARC_CREDITO` REAL,
  `VL_REC_PROPRIO` REAL,
  `VL_PERC_RISCO_STN` REAL,
  `VL_PERC_RISCO_FUNDO_CONST` REAL,
  `VL_REC_PROPRIO_SRV` REAL,
  `VL_AREA_FINANC` REAL,
  `CD_SUBPROGRAMA` INTEGER,
  `VL_PRODUTIV_OBTIDA` REAL,
  `DT_FIM_COLHEITA` TEXT,
  `DT_FIM_PLANTIO` TEXT,
  `DT_INIC_COLHEITA` TEXT,
  `DT_INIC_PLANTIO` TEXT,
  `VL_JUROS_ENC_FINAN_POSFIX` REAL,
  `VL_PERC_CUSTO_EFET_TOTAL` REAL,
  `CD_CNPJ_CADASTRANTE` INTEGER,
  `VL_AREA_INFORMADA` REAL,
  `CD_CICLO_CULTIVAR` INTEGER,
  `CD_TIPO_SOLO` INTEGER,
  `PC_BONUS_CAR` INTEGER,
   PRIMARY KEY (REF_BACEN, NU_ORDEM)
);"
)

dbWriteTable(conn, "sicor_operacao_basica", sicor_operacao_basica, append = TRUE)

# Write sicor_liberacao_recursos data to DB ----
dbExecute(
  conn,
  "CREATE TABLE `sicor_liberacao_recursos` (
  `VL_LIBERADO` REAL,
  `REF_BACEN` INTEGER,
  `NU_ORDEM` INTEGER,
  `year_treat` TEXT,
   PRIMARY KEY (year_treat, REF_BACEN, NU_ORDEM),
   FOREIGN KEY (REF_BACEN, NU_ORDEM) REFERENCES sicor_operacao_basica(REF_BACEN, NU_ORDEM)
);"
)

dbWriteTable(conn, "sicor_liberacao_recursos", sicor_liberacao_recursos, append = TRUE)


# Write CAR data to DB ----
dbExecute(
  conn,
  "CREATE TABLE `car_aml` (
  `cod_imovel` TEXT PRIMARY KEY,
  `num_area` REAL,
  `municipio` TEXT,
  `cod_estado` TEXT,
  `geometry` geometry,
  `wkt` TEXT);"
)


dbWriteTable(conn, "car_aml", car_aml_df, append = TRUE)
dbExecute(conn, "SELECT AddGeometryColumn('car_aml', 'geom', 4674, 'POLYGON', 'XY');")
dbExecute(conn, "UPDATE car_aml SET geom = ST_GeomFromText(wkt, 4674);")

# Write PRODES data to DB ----
dbExecute(
  conn,
  "CREATE TABLE `prodes` (
  `fid` INTEGER PRIMARY KEY ,
  `state` TEXT,
  `path_row` TEXT,
  `main_class` TEXT,
  `class_name` TEXT,
  `sub_class` TEXT,
  `def_cloud` REAL,
  `julian_day` INTEGER,
  `image_date` REAL,
  `year` REAL,
  `area_km` REAL,
  `scene_id` INTEGER,
  `source` TEXT,
  `satellite` TEXT,
  `sensor` TEXT,
  `uuid` TEXT,
  `wkt` TEXT,
  `_ogr_geometry_` geometry);"
)

prodes_df <- as.data.frame(prodes)
prodes_df$wkt <- prodes@ptr$getGeometryWKT() # create wkt to load on Spatialite
dbWriteTable(conn, "prodes", prodes_df, append = TRUE)
dbExecute(conn, "SELECT AddGeometryColumn('prodes', 'geom', 4674, 'MULTIPOLYGON', 'XY');")
dbExecute(conn, "UPDATE prodes SET geom = ST_GeomFromText(wkt, 4674);")

# Write sicor_propriedades data to DB ----
dbExecute(
  conn,
  "CREATE TABLE `sicor_propriedades` (
  `REF_BACEN` INTEGER,
  `NU_ORDEM` INTEGER,
  `cod_imovel` TEXT,
  `uf` TEXT,
  FOREIGN KEY (REF_BACEN, NU_ORDEM) REFERENCES sicor_operacao_basica(REF_BACEN, NU_ORDEM),
  FOREIGN KEY (cod_imovel) REFERENCES car_aml(cod_imovel)
);"
)

dbWriteTable(conn, "sicor_propriedades", sicor_propriedades, append = TRUE)

# Write deforestation and loan values (filtred) data to DB ----
dbExecute(
  conn,
  "CREATE TABLE `deforestation_and_values_f` (
  `cod_imovel` TEXT PRIMARY KEY,
  `deforestation_ha` REAL,
  `municipio` TEXT,
  `cod_estado` TEXT,
  `area_forest_2016` REAL,
  `vl_liberado` REAL
);"
)

dbWriteTable(conn, "deforestation_and_values_f", deforestation_and_values_filtered, append = TRUE)


# Close the database connection
dbDisconnect(conn)
```
