---
title: "Rural Credit and Deforestation in Brazilian Amazon Forest"
subtitle: GEOG 514B - Topics in Environmental Geography - Dr. Jessica Dempsey
author: George Porto Ferreira
date: "`r format(Sys.time(), '%Y-%B-%d')`"
bibliography: inpe2.bib
linenumbers: false
numbersections: FALSE
output:
  pdf_document:
    toc: yes
---

```{r setup, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(dplyr)
library(sf)
library(fs)
library(tidyverse)
library(httr)

options(timeout = 1500) # (in seconds) 
#if you are in a limited connection, increase the timeout

```

# Central Bank of Brazil - SICOR Database

The Central Bank of Brazil makes rural credit data available from the SICOR Database. More details about the types of credit and other information can be accessed through the link <https://www.bcb.gov.br/estabilidadefinanceira/creditorural>

In this work, we accessed data from two tables with data on rural properties and loan amounts received (SICOR_PROPRIEDADES & SICOR_LIBERACAO_RECURSOS)

The model and data dictionary are available at the link below: <https://www.bcb.gov.br/estabilidadefinanceira/creditorural?modalAberto=tabelas_sicor>

The code below loads the data from these two tables. They will be downloaded if they do not exist in the working directory.

```{r SICOR tables, eval=FALSE}

# https://www.bcb.gov.br/htms/sicor/DadosBrutos/SICOR_PROPRIEDADES.gz
# https://www.bcb.gov.br/htms/sicor/DadosBrutos/SICOR_LIBERACAO_RECURSOS.gz

url_sicor <- 'https://www.bcb.gov.br/htms/sicor/DadosBrutos/'

if (file.exists(here::here("SICOR_PROPRIEDADES.gz"))) {
  sicor_propriedades <- gzfile(here::here("SICOR_PROPRIEDADES.gz")) |>
  read.csv(sep = ";")  
  } else {
    download.file(paste0(url_sicor,'SICOR_PROPRIEDADES.gz'), 
                  here::here("SICOR_PROPRIEDADES.gz"))
    sicor_propriedades <- gzfile(here::here("SICOR_PROPRIEDADES.gz")) |> 
      read.csv(sep = ";")
  }

sicor_propriedades <- mutate(
  sicor_propriedades, ref_banuor = paste0(X.REF_BACEN, NU_ORDEM)
  ) |>
  dplyr::select(-X.REF_BACEN, -NU_ORDEM) |>
  dplyr::filter(substr(CD_CAR,1,2) %in%  c('RO','PA','MT'))


if (file.exists(here::here("SICOR_LIBERACAO_RECURSOS.gz"))) {
  sicor_liberacao_recursos <- gzfile(here::here("SICOR_LIBERACAO_RECURSOS.gz")) 
  |>
  read.csv(sep = ";")  
  } else {
    download.file(paste0(url_sicor,'SICOR_LIBERACAO_RECURSOS.gz'),
                  here::here("SICOR_LIBERACAO_RECURSOS.gz"))
    sicor_liberacao_recursos <- gzfile(here::here("SICOR_LIBERACAO_RECURSOS.gz")) 
    |> 
    read.csv(sep = ";")
  }

sicor_liberacao_recursos <- mutate(
  sicor_liberacao_recursos, year = substr(X.LIR_DT_LIBERACAO,7,10), 
                                   ref_banuor = paste0(REF_BACEN, NU_ORDEM)) |>
  dplyr::select(-X.LIR_DT_LIBERACAO, -REF_BACEN, -NU_ORDEM)|>
  dplyr::filter(year > 2012 & year < 2023 ) |>
  group_by(ref_banuor)|>
    summarize(total = sum(LIR_VL_LIBERADO))|>
  ungroup()

```

# Brazilian Forestry Service - Rural Environmental Registry Database

The Brazilian Forestry Service makes spatial data with the delimitation of rural properties available. Public access offers only the manual download municipality by municipality through the link <https://www.car.gov.br/publico/municipios/downloads>

The Legal Amazon comprises 9 Brazilian states. However, in this study only, the shapefiles of rural properties in Rondônia, Mato Grosso and Pará were downloaded.

Before running the code below, the shapefiles for each state must be downloaded manualy and stored in a folder named "car" under your "Downloads" folder.

The code below concatenates the shapefiles from the 141 municipalities in the state of Mato Grosso, the 144 municipalities in Pará, and the 52 municipalities in the state of Rondônia.

```{r CAR, eval=FALSE}

# Unzip and rename Rural Properties Windows environment------------

downloaded <- paste0("C:/Users/",Sys.getenv("USERNAME"),"/Downloads/car")
ext <- c(".dbf", ".shp", ".shx",".prj" )

zipfiles <- downloaded %>%
  dir_ls()

for (i in 1:length(zipfiles)){
unzip(zipfiles[i], exdir = substr(zipfiles[i], 1, nchar(zipfiles[i])-4))
}

level1_unziped_folders <- substr(zipfiles, 1, nchar(zipfiles)-4) %>% 
  dir_ls() %>%
  str_subset('AREA_IMOVEL.zip$')

for (i in 1:length(level1_unziped_folders)){
  unzip(
    level1_unziped_folders[i], exdir = substr(zipfiles[i], 1, nchar(zipfiles[i])-4)
    )
  file.rename(
    paste0(substr(zipfiles[i], 1, nchar(zipfiles[i])-4),'/AREA_IMOVEL',ext),
    paste0(substr(zipfiles[i], 1, nchar(zipfiles[i])-4),
           substr(zipfiles[i], nchar(zipfiles[i])-17, nchar(zipfiles[i])-4),ext)
    )
  }

# Getting all file paths ------------

shapefiles <- downloaded %>%
  dir_ls(recurse = TRUE) %>%
  str_subset('.shp$')

# Loading all files  -----------------
car_pamtro <- shapefiles %>%
  map(st_read) %>%
  bind_rows() %>%
  st_transform(crs = 5880)
```

# National Institute for Space Research - PRODES Database

Every year the National Institute for Space Research (INPE) maps out the increase in deforestation based on satellite images. Information about the PRODES Project can be accessed through the link [http://www.obt.inpe.br/OBT/assuntos/programas/amazonia/prodes](#0){.uri}. The methodology [@shimabukuroMappingMonitoringDeforestation2000] [@valerianoMonitoringTropicalForest2004] used since 1988 has been improved over the years, and today, all policies to combat deforestation and the Brazilian NDC use PRODES data as a basis.

The code below loads the data for processing.

```{r Deforestation Data PRODES, eval=FALSE}

prodes_path <- paste0("C:/Users/",Sys.getenv("USERNAME"),"/Downloads/")

prodes_url <- 
  'http://terrabrasilis.dpi.inpe.br/download/dataset/legal-amz-prodes/vector/yearly_deforestation.zip'
  
if (file.exists(paste0(prodes_path,'yearly_deforestation.shp'))) {
    prodes <- st_read(paste0(prodes_path,'yearly_deforestation.shp'), 
     query = 
      "SELECT * FROM yearly_deforestation WHERE year > 2008 AND state IN('RO', 'PA', 'MT')"
      )|>
    st_transform(crs = 5880)
  } else {
    download.file(prodes_url, paste0(prodes_path,'yearly_deforestation.zip'))
    unzip(paste0(prodes_path,'yearly_deforestation.zip'), exdir = prodes_path)
    prodes <- st_read(paste0(prodes_path,'yearly_deforestation.shp'),
     query = 
      "SELECT * FROM yearly_deforestation WHERE year > 2008 AND state IN('RO', 'PA', 'MT')"
      )|>
    st_transform(crs = 5880)
    file.remove(paste0(prodes_path,'yearly_deforestation.zip'))
}

```

# Data analysis

The analysis starts by identifying which rural properties suffered any deforestation after August 2008. It is essential to understand that the PRODES data always comprise 12 months from August of one year to July of the following year.

Another critical concept to be understood is the expiration of the punitive intention. According to law 9605, in general, deforestation that has been committed five years or more ago can no longer receive a fine. In addition, law 12651 determines that deforestation committed without authorization after July 22, 2008, is not eligible for regularization.

For these reasons, only data from PRODES after July 2008 are being used in this analysis. Furthermore, the SICOR data began in 2013, which allows us to verify if, after 2013, any properties that received rural credit had committed deforestation in the previous five years.

```{r Calculations, eval=FALSE}

# WARNING! There will be duplicated values for deforestation areas due to CAR overlaps ----
deforestation_in_car <- st_intersection(
  st_make_valid(car_pamtro), st_make_valid(prodes)
  ) %>% 
  mutate(area = st_area(.$geometry), 
         CD_CAR = paste0(substr(COD_IMOVEL,1,2), 
                         substr(COD_IMOVEL,4,10), 
                         substr(COD_IMOVEL,12,43)
                         )
         ) %>% 
  dplyr::select(NUM_AREA:NOM_MUNICI, TIPO_IMOVE, image_date, year, area, CD_CAR) %>%
  dplyr::group_by(CD_CAR) %>% 
    dplyr::summarise(deforestation = sum(area)) %>% 
  ungroup()

# WARNING! There will be duplicated values for deforestation areas ----
# inherited from deforestation_in_car and also duplicated values in
# credit values because of cardinality between liberacao_recursos and 
# propriedades is many-to-many
deforestation_and_values <- full_join(
  deforestation_in_car, sicor_propriedades, by = "CD_CAR"
  ) |>
  full_join(sicor_liberacao_recursos, by = "ref_banuor") |>
  mutate(uf = substr(CD_CAR,1,2))

list_car_w_credit <- dplyr::filter(deforestation_and_values, total >0)

# Resolving CAR overlaps to calculate Deforestation in CAR with Credit----
prodes_w_credit_RO <- dplyr::filter(list_car_w_credit, uf == 'RO') |> 
  st_union() |>
  st_area()

prodes_w_credit_MT<- dplyr::filter(list_car_w_credit, uf == 'MT') |> 
  st_union() |>
  st_area()

prodes_w_credit_PA<- dplyr::filter(list_car_w_credit, uf == 'PA') |> 
  st_union() |>
  st_area()


# Resolving duplicated values in credit with deforestation
credit_in_car_w_prodes <- dplyr::filter(
  deforestation_and_values, as.numeric(deforestation) > 0
  )|>
  as.data.frame() |>
  group_by(ref_banuor)|>
    summarize(credit = first(total), 
              qtd_car = n(),
              uf = first(uf)) |>
  ungroup()

# Credit in CAR with deforestation
credit_PA_w_prodes<- dplyr::filter(credit_in_car_w_prodes, uf == 'PA') 
credit_MT_w_prodes<- dplyr::filter(credit_in_car_w_prodes, uf == 'MT') 
credit_RO_w_prodes<- dplyr::filter(credit_in_car_w_prodes, uf == 'RO')

sum(credit_PA_w_prodes$credit, na.rm = TRUE)
sum(credit_MT_w_prodes$credit, na.rm = TRUE)
sum(credit_RO_w_prodes$credit, na.rm = TRUE)

# Resolving duplicated values in credit without deforestation
credit_in_car_wo_prodes <- dplyr::filter(
  deforestation_and_values, as.numeric(deforestation) == 0 | is.na(deforestation)
  )|>
  as.data.frame() |>
  group_by(ref_banuor)|>
    summarize(credit = first(total), 
              qtd_car = n(),
              uf = first(uf)) |>
  ungroup()

# Credit in CAR without deforestation
credit_PA_wo_prodes<- dplyr::filter(credit_in_car_wo_prodes, uf == 'PA') 
credit_MT_wo_prodes<- dplyr::filter(credit_in_car_wo_prodes, uf == 'MT') 
credit_RO_wo_prodes<- dplyr::filter(credit_in_car_wo_prodes, uf == 'RO')

sum(credit_PA_wo_prodes$credit, na.rm = TRUE)
sum(credit_MT_wo_prodes$credit, na.rm = TRUE)
sum(credit_RO_wo_prodes$credit, na.rm = TRUE)
```

# References {.unnumbered}

::: {#refs}
:::
